\LARGE \bf Mean Estimation from Single-bit Measurements

### 2017-09-26 Writeups/onebit_full_V1.tex
Add files via upload

\usepackage{mathptmx} 
\usepackage{times} 
\usepackage{amsmath} 
\usepackage{amsbsy} 
\usepackage{amssymb}
%\usepackage{newtxtext, newtxmath}
\usepackage{mathrsfs}
\usepackage{comment}
\usepackage[export]{adjustbox}
\usepackage{tikz}
\usetikzlibrary{external,positioning,decorations.pathreplacing,shapes,arrows,patterns}

@@

\newcommand{\mmse}{\mathsf{mmse}}
\newcommand{\supp}{\mathrm{supp} }
\renewcommand\vec[1]{\ensuremath\boldsymbol{#1}}
\newenvironment{proof}{\paragraph*{Proof}}{\hfill$\square$ \newline}
\newcommand{\sgn}{\mathrm{sgn} }
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand*{\QEDA}{\hfill\ensuremath{\square}}

@@

\begin{document}
\graphicspath{{../Figures/}}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

@@

\subsection*{Related Works}
%\subsubsection*{Statistical inference under communication constraints}

@@

@@

The first main results of this paper, as described in Theorem~\ref{thm:adpative_lower_bound} below, states that the asymptotic relative efficiency of and adaptive estimator cannot get below $\pi/2$. Next, we provide a particular adaptive estimation scheme and show in Theorem~\ref{thm:sgd} that asymptotically attains this efficiency with equality. Finally, in Theorem~\ref{thm:opt_one_step} we provide an adaptive estimation scheme that is optimal in the sense that at each step $i$, the message $M_i$ that minimizes the MSE given $X_i$ and the previous $M^{i-1}$ messages is chosen. 
%While we provide expressions for the optimal message, evaluating the asymptotic efficiency of this estimation scheme turns out to be a challenging task. We therefore leave open the question whether this scheme attains the lower bound with equality, and only provide numerical result that support this conjecture.  \\

@@

\subsection{A Lower Bound on Adaptive Single-Bit Schemes}
Our first results asserts that the relative efficiency \eqref{eq:relative_efficiency} of any adaptive estimation scheme is bounded from below by $\pi/2$, as follows from the following theorem:
\begin{thm}[minimal relative effeciency] \label{thm:adpative_lower_bound}
Let $\widehat{\theta}_n$ be any estimator of $\theta$ in the adaptive setting of Fig.~\ref{fig:sequential}.
% Let $\pi(\theta)$ be an absolutely continuous prior distribution on $\Theta$ with density function converges to zero at the endpoints of the interval $\Theta$. 
Assumes that $P(d\theta)$ converges to zero at the endpoints of the interval $\Theta$. Then
\[
\mathbb E (\theta-\theta_n)^2 \geq  \frac{\pi \sigma^2 }{2n +\pi \sigma^2  I_0}  =  \frac{\pi}{2n}\sigma^2+o(n^{-1}),
\]
where 
\[
I_0 = \mathbb E \left( \frac{d}{d\theta} \log P_\theta(\theta) \right)^2
\]
is the Fisher information with respect to a location model in $\theta$. 
\end{thm}

\subsubsection*{Sketch of Proof}
The main idea in the proof is to bound from above the Fisher information of any set of $n$ single-bit messages with respect to $\theta$. Once this bound is achieved, the result follows by using the Van-Trees inequality \cite{gill1995applications} which bounds from below the MSE of any estimator of $\theta$ by the inverse of the expected value of the aforementioned Fisher information plus $I_0$. The details are given in the Appendix.\\

Next, we present an adaptive estimation scheme which attains the minimal asymptotic efficiency $\pi/2$. 
%Theorem~\ref{thm:adpative_lower_bound}. This estimation scheme can be seen as a gradient descent

\subsection{Asymptotically Optimal Estimator}
Consider the following estimator $\widehat{\theta}_n$ for $\theta$:  set 
\begin{equation}
\label{eq:sgd_alg}
\theta_n = \theta_{n-1} +  \gamma_n \sgn (X_n - \theta_n), \quad n = 1,2,\ldots,
\end{equation}
where $\left\{\gamma_n \right\}_{n=1}^\infty$ is any strictly positive sequence satisfying 
\begin{enumerate}
\item[(i)] $\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n)$ \\
\item[(ii)] $\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty$ 
for some $0< \lambda \leq 1$.
\end{enumerate}
(e.g. $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). The $n$th step estimation is defined by 
\begin{equation} \label{eq:sgd_est}
\widehat{\theta}_n =  \frac{1}{n} \sum_{i=1}^n  \theta_i. 
\end{equation}

As the following theorem shows, the estimator $\widehat{\theta}_n$ defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg} attains the minimal asymptotic efficiency as established by Theorem~\ref{thm:adpative_lower_bound}.
\begin{thm} \label{thm:sgd}
The sequence $\widehat{\theta}_n$ of \eqref{eq:sgd_alg} satisfies
\[
\sqrt{n} \left( \widehat{\theta}_n - \theta \right) \overset{d}{\rightarrow} \mathcal N \left(0,  \pi \sigma^2 /2 \right).
\]
\end{thm}

@@

Note that $\theta_0$ is not explicitly defined in 
in equation \eqref{eq:sgd_est}. Arguably, a reasonable initialization is $\theta_0 = \mathbb E [\theta]$. However, Theorem~\ref{thm:sgd} implies that the asymptotic behavior of the estimator, is indifferent to this initialization. Thus, the optimal efficiency is attained without using the prior distribution on $\theta$. Nevertheless, the bound in Theorem~\ref{thm:adpative_lower_bound} suggests that the estimation error can be greatly reduced whenever the prior of $\theta$ provides a lot of information $I_0$ on its location. In order to incorporate this information in a more meaningful way, we present a Bayesian estimation scheme that is conjectured to attain the optimal asymptotic efficiency.

\subsection{One-step Optimal Estimation}
In this subsection we present an estimation scheme that posses the property of \emph{one-step optimality}. That is, at each step $i$, the $i$th encoder designs the detection region $M_i^{-1}(1)$ such that the MSE given $M^i$ is minimal. In other word, this scheme designs the messages in a greedy manner, such that the MSE at step $i$ is minimal given the current state of the estimation described by $M^{i-1}$. \\

The structure of the message that minimizes the next step MSE is given by the following theorem:
\begin{thm}[optimal one-step estimation] \label{thm:opt_one_step}
Let $\pi(\theta)$ be an absolutely continuous log-concave probability distribution. Given a sample $X$ from the distribution $\mathcal N(\theta, \sigma^2)$, define 
\begin{equation}
\label{eq:adaptive_main_message}
M = \sgn(X - \tau),
\end{equation}
where $\tau$ satisfies the equation
\begin{equation}
 \label{eq:fixed_point}
 \tau = \frac{m^-(\tau) + m^+(\tau)}{2},
\end{equation}
with
\begin{align*}
m^-(\tau)  & = \frac{\int_{-\infty}^{\tau} \theta P(d\theta) }{\int_{-\infty}^{\tau} P(d\theta)} ,\\
m^+(\tau) & = \frac{\int_{\tau}^\infty \theta P(d\theta) }{\int_{\tau}^\infty P(d\theta)} .
\end{align*}
Then for any estimator $\widehat{\theta}$ which is a function of $M'(X) \in \{-1,1\}$, we have
\begin{equation}
\label{eq:opt_cond}
\mathbb E \left(\theta-\widehat{\theta}(M')\right)^2 \geq  \mathbb E \left(\theta- \mathbb E[\theta|M]\right)^2,
\end{equation}
\end{thm}

\begin{proof}
The proof is completed by the following two lemmas, proofs of which can be found in the Appendix:
\begin{lem} \label{lem:unique}
Let $f(x)$ be a log-concave probability density function. Then the equation 
\begin{equation}
\label{eq:lem_fixed_point}
2x = \frac{\int_x^\infty uf(u)du}{\int_x^\infty f(u)du} + \frac{\int_{-\infty}^x uf(u)du}{\int_{-\infty}^x f(u)du} 
\end{equation}
has a unique solution.
\end{lem}
\end{proof}

\begin{lem} \label{lem:adaptive}
Let $U$ be an absolutely continuous random variable with
pdf $P(du)$. Then the one-bit message $M^\star$ that minimizes
\[
\int \left( u - \mathbb E[U|M(u)]  \right)^2 P(du)
\]
is given by
\[
M^\star  =  \sgn(U - \tau),
\]
where $\tau$ is the unique solution to the equation
 \[
2 \tau = \frac{\int_{\tau}^\infty u P(du)} {\int_{\tau}^\infty P(du)} + \frac{\int_{-\infty}^{\tau} u P(du)}{\int_{-\infty}^{\tau} P(du)}.
\]
\end{lem}

Theorem~\ref{thm:opt_one_step} suggests the following Bayesian estimation scheme: set $P_0(t) = \pi(\theta)$, and for $n\geq 1$ update the prior as
\begin{align}
P_n(t) = & P(\theta=t |M^n) = \frac{ P\left( \theta=t | M^{n-1} \right) P(M_n | \theta = t , M^{n-1})  } { P(M_n | M^{n-1} )} \nonumber \\ 
& = \alpha_n  P_{n-1}(t) \Phi\left(M_n \frac{ t - \tau_{n-1} }{\sigma} \right), \label{eq:density_update}
\end{align}
where $\alpha_n$ is a normalization coefficient that equals to
\[
\alpha_n = \left(\int_{\mathbb R} P_{n-1}(t) \Phi\left(M_n \frac{t- \tau_{n-1} }{\sigma} \right)  dt \right)^{-1}. 
\]
The $n$th estimate for $\theta$ is the conditional expectation of $\theta$ given $M^n$, namely
\begin{equation}
\theta_n = \mathbb E \left[ \theta| M^n\right] = \int_{-\infty}^\infty t P_n(t) dt. \label{eq:estimator_update}
\end{equation}
Next, solve equation \eqref{eq:fixed_point} with the updated prior $P_n(t)$ instead of $P(\theta)$. Note that since the standard normal cdf $\Phi(x)$ is log-concave, the updated prior $P_n(t)$ remains log-concave and thus a unique solution to \eqref{eq:fixed_point} is guaranteed by Lemma~\ref{lem:unique}. Finally, update the $n+1$th message as
\begin{equation}\label{eq:message_update}
M_{n+1} = \sgn(X_{n+1}-\tau_n).
\end{equation}

Since the equation \eqref{eq:fixed_point} has no analytic solution in general, it is hard to derive the asymptotic behavior of the Bayes estimator defined by \eqref{eq:estimator_update} and \eqref{eq:message_update}. We conjecture, however, that it attains the asymptotic relative efficiency of $\sigma^2\pi/2$, as can be observed from the numerical simulation illustrated in Fig.~\ref{fig:adaptive_error}. Also shown in Fig.~\ref{fig:adaptive_error} are the normalized MSE of the estimator defined by \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, as well as the MSE achieved by the sample mean for the same sample realization.

 % In fact, the numerical simulation in Fig.~\ref{fig:simulation} shows that this scheme outperform the estimator of \eqref{eq:sgd_est} even when the prior on $\Theta$ is least favorable for Bayes estimation \cite{Bickel1981}.

\begin{figure}
\begin{center}
\begin{tikzpicture}
\node at (0,0) {\includegraphics[scale=0.4]{error}};
\end{tikzpicture}
\caption{Normalized empirical risk $n\left(\widehat{\theta}_n-\theta\right)^2$ versus number of samples $n$ for $100$ Monte Carlo trials, where $\theta$ is chosen uniformly in the interval $(-5,5)$. \label{fig:adaptive_error}  }
\end{center}
\end{figure}

@@

\subsection{Threshold Detection}
We assume that each message is of the form
\[
M_i = \sgn(t_i - X_i) = \begin{cases} 1 & X_i< t_i, \\
-1 & X_i > t_i,
\end{cases}  
\]
where $t_i\in\mathbb R$ is the \emph{threshold} of the $i$th detector applied by the encoder on it sample $X_i$. In order to characterize the asymptotic behavior of estimation from these sequence of messages, we consider the \emph{density} of $n$ threshold values defined as:
\[
\lambda_n([a,b]) = \frac{1}{n} \left| \mathcal T \cap [a,b] \right|,
\]
where $\mathcal T  = \left\{t_n, n \in \mathbb N \right\}$ is the collection of all threshold values. We further assume that $\lambda_n$ converges (weakly) to a probability measure $\lambda(t)$ on $\mathbb R$.  \par
In order to estimate $\theta$ from the messages $M^n$, we consider the maximum likelihood (ML) estimator. The log-likelihood function of $\theta$ is given by
\[
l(M^n  |\theta ) =  \sum_{i=1}^n \log \left( \Phi \left( M_i \frac{t_i - \theta}{\sigma} \right) \right). 
\]
Since $\Phi(x)$ is a log concave function, the log-likelihood function has a unique maximizer $\widehat{\theta}_n$ which is the ML estimator.  \\

Next, we show that under the assumptions above, the sequence of messages $M^n$ defines a local asymptotic normal (LAN) family of probability distributions. Therefore, the MSE in estimating $\theta$ from the messages $M^n$ satisfies a local asymptotic minimax property with respect to the \emph{precision} parameter of this LAN family \cite{van2000asymptotic}. Moreover, we conclude the ML estimator is locally asymptotic minimax.

@@

>>>>

\usepackage{mathptmx} 
\usepackage{times}

\usepackage{amsmath} 
\usepackage{amsbsy} 
\usepackage{amssymb}
%\usepackage{newtxtext, newtxmath}
\usepackage{mathrsfs}
\usepackage{comment}
\usepackage[export]{adjustbox}
\usepackage{tikz}
\usetikzlibrary{external,positioning,decorations.pathreplacing,shapes,arrows,patterns}

@@

\newcommand{\mmse}{\mathsf{mmse}}
\newcommand{\unif}{\mathsf{unif}}
\newcommand{\supp}{\mathrm{supp} }
\renewcommand\vec[1]{\ensuremath\boldsymbol{#1}}
\newenvironment{proof}{\paragraph*{Proof}}{\hfill$\square$ \newline}
\newcommand{\sgn}{\mathrm{sgn} }
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand*{\QEDA}{\hfill\ensuremath{\square}}

@@

\begin{document}
\graphicspath{{../Figs/}}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

@@

\subsection*{Related Works}
%\subsubsection*{Statistical inference under communication constraints}
{\color{red} Shamir \cite{Shamir} showed that in hide-and-seek hypothesis testing problem, bitrate constraint may lead to a significant reduction in detection performance compared to the unrestricted case. This result is then applied to online learning, stochastic optimization and sparse PCA. Shamir's results shows that the penalty is in the dimension ($d/R$ instead of $lod(d)$). In our setting the dimension is $1$.} \\

@@

\section{Relation to Remote Multiterminal Source Coding \label{sec:ceo}}
In this section we draw a connection between our general problem of mean estimation from one-bit samples, to the remote multiterminal source coding problem \cite{berger1996ceo}, also known as the CEO problem. The setting of the CEO includes $n$ encoders, each has access to a noisy version of a random source sequence. 
% $\left\{ U_j,\,j=1,2,\ldots \right\}$. 
The $i$th encoder observes $k$ noisy source symbols and transmit $R_i k$ bits to a central estimator. \par
Assuming that $\theta$ is drawn randomly once from a prior $\pi(\theta)$, mean estimation from one-bit under distributed encoding (setting (iii) in the Introduction) corresponds to the CEO setting with block length $k=1$, where the $i$th encoder observes
\begin{equation}
\label{eq:Gaussian_channel}
X_i = \theta + \sigma Z_i,
\end{equation}
with $Z_1,\ldots,Z_n$ i.i.d. standard normal, and uses $R_i=1$ bits to transmit its observation. As a result, the quadratic distortion in the optimal source coding scheme for the CEO with $n$ terminals at rates $R_1 = \ldots = R_n = 1$ and Gaussian observation noise at each estimate of variance $\sigma^2$, provides a lower bound on the MSE distortion in estimating $\theta$ in the distributed encoding setting. 
 \\

A closed-form expression for the MSE distortion in the CEO is known only for the case where the source sequence is sampled independently from the normal distribution \cite{prabhakaran2004rate}. By using the characterization of this minimal distortion as the number of terminals goes to infinity, we conclude the following:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\pi(\theta) = \mathcal N(0,\sigma_\theta^2)$. Then any estimator $\widehat{\theta}_n$ of $\theta$ in 
the distributed setting satisfies
\begin{equation} \label{eq:ceo_bound}
 n\mathbb E \left( \theta - \theta_n \right)^2 \geq \frac{4\sigma^2}{3} + o(1).
\end{equation}
\end{prop}

\begin{proof}
We consider the expression \cite[Eq. 10]{chen2004upper} that provides the minimal distortion $D^\star$ in the CEO with $L$ observers and under a total sum-rate $R_\Sigma = R_1 + \ldots +R_L$:
\begin{equation} \label{eq:ceo_optimal_sumrate}
R_{\Sigma} = \frac{1}{2} \log^+ \left[ \frac{\sigma_\theta^2}{D^\star} \left( \frac{D^\star L}{ D^\star L - \sigma^2 + D^\star \sigma^2 / \sigma_\theta^2 }\right)^L  \right].
\end{equation}
Assuming $R_\Sigma = n$ and $L=n$, we get
\begin{equation} \label{eq:ceo_optimal_sumrate}
n = \frac{1}{2} \log \left[ \frac{\sigma_\theta^2}{D^\star} \left(\frac{ D^\star n }{D^\star n - \sigma^2 + D^\star \sigma^2/\sigma_\theta^2 }  \right)^n  \right].
\end{equation}
The value of $D^\star$ that satisfies the equation above describes the MSE under an optimal allocation of the sum-rate $R_\Sigma = n$ among the $n$ encoders. Therefore, $D^\star$ provides a lower bound to the CEO distortion with $R_1=\ldots,R_n = 1$ and hence a lower bound to the minimal MSE in estimating $\theta$ in the distributed encoding setting. By considering $D^\star$ in \eqref{eq:ceo_optimal_sumrate} as $n\rightarrow \infty$, we conclude that 
\[
D^\star = \frac{ \frac{4\sigma^2}{3} }{n + \frac{4 \sigma^2}{3 \sigma_\theta^2} } + O(e^{-n}) =  \frac{4\sigma^2}{3n} + o(n^{-1}). 
\]
\end{proof}
Since the lower bound \eqref{eq:ceo_bound}  was derived assuming optimal allocation of the $n$ bits among the encoders, it may seem as if a tighter bound can be obtained in our case by considering the CEO distortion with $R_1=\ldots,R_n = 1$. However, an upper bound for the CEO distortion under this condition follows from \cite[Prop. 5.2]{KipnisRini2017}, and leads to
\[
D_{CEO} \leq  \left( \frac{1}{\sigma_\theta^2} +  \frac{3n}{4\sigma^2 + \sigma_\theta^2} \right)^{-1}   =
\frac{4 \sigma^2}{3n} +  \frac{\sigma_\theta^2}{3n} + o(n^{-1}),
\]
which is equivalent to \eqref{eq:ceo_bound} when $\sigma_\theta$ is small. \par
We conclude therefore that the difference between the MSE lower bound \eqref{eq:ceo_bound} and the actual MSE in the distributed encoding setting, is attributed exclusively to the ability to perform coding over blocks. Namely, the ability to consider $k>1$ independent realizations of $\theta$ versus only one in ours. In other words, it is the ability to exploit the geometry of a high-dimensional product space, rather than the distributed nature of the problem, that distinguishes between the CEO distortion and the mean estimation from one-bit samples.\par
In the next section we show that the ARE in adaptive encoding setting does not exceeds $\pi/2$, and thus provides a tighter lower bound for the distributed encoding setting than $4/3$ of  \eqref{eq:ceo_bound}.

\section{Relation to Decentralized Detection}
Once the decision rule of each sensor or encoder is fixed, the optimal estimation of $\theta$ is determined using the maximum a posteriori probability rule. Hence, the distributed setting is reduced to finding the optimal decision rule of each encoder. When the parameter space $\Theta$ is finite, the characterization of the optimal decision rules was considered by Tsitsiklist in \cite{tsitsiklis1988decentralized}. It was shown there that if the cardinality of $\Theta$ is at most $M$ and the probability of error criterion is used, than no more than $M(M-1)/2$ different decision rules are necessary in order to attain probability of error decreasing exponentially with the optimal exponent. That is, some beyond $M(M-1)/2$ some decision rules can be repeated. A version of this problem for the adaptive scenario was also considered in \cite{5751320}, where it was shown that with a specific two-stages feedback, asymptotically there is no gain in feedback compared to the fully distributed setting. 
 \\
Wei et. al. \cite{904560} considered the case where a single threshold is applied to all quantizers, and showed how to chose the optimal threshold. ARE and minimax ARE of estimation from distributed quantized measurements that is based on Fisher information was considered in \cite{4244748, 6882252, chen2010performance, 5184907}. However, the error criterion considered in these works is the Fisher information rather than a risk or ARE criterion. Moreover, these works provide optimality conditions only under the assumption the the same quantizer is used at each encoder.

@@

The first main results of this paper, as described in Thm.~\ref{thm:adpative_lower_bound} below, states that the ARE of any adaptive estimator cannot be lower than $\pi/2$. Next, we provide a particular adaptive estimation scheme and show in Thm.~\ref{thm:sgd} that its efficiency is $\pi/2$. Finally, in Thm.~\ref{thm:opt_one_step}, we provide an adaptive estimation scheme that is one-step optimal in the sense that at each step $i$, the message $M_i$ that minimizes the MSE given $X_i$ and the previous $M^{i-1}$ messages is chosen. While it is not clear whether the efficiency of this last scheme is $\pi/2$, numerical simulations suggests that the MSE of this scheme times $n$ also converges to $\pi/2$.

@@

\subsection{A lower bound on adaptive one-bit schemes}
Our first results asserts that the ARE \eqref{eq:relative_efficiency} of any adaptive estimation scheme is bounded from below by $\pi/2$, as follows from the following theorem:
\begin{thm}[minimal relative effeciency] \label{thm:adpative_lower_bound}
Let $\widehat{\theta}_n$ be any estimator of $\theta$ in the adaptive setting of Fig.~\ref{fig:sequential}.
% Let $\pi(\theta)$ be an absolutely continuous prior distribution on $\Theta$ with density function converges to zero at the endpoints of the interval $\Theta$. 
Assumes that $\pi(\theta)$ converges to zero at the endpoints of the interval $\Theta$. Then
\[
\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{\pi \sigma^2 }{2n +\pi \sigma^2  I_0}  =  \frac{\pi}{2n}\sigma^2+o(n^{-1}),
\]
where 
\[
I_0 = \mathbb E \left( \frac{d}{d\theta} \log \pi (\theta) \right)^2
\]
is the Fisher information with respect to a location model in $\theta$. 
\end{thm}

\subsubsection*{Sketch of Proof}
The main idea in the proof is to bound from above the Fisher information of any set of $n$ single-bit messages with respect to $\theta$. Once this bound is achieved, the result follows by using the van-Trees inequality \cite[Thm. 2.13]{tsybakov2008introduction},\cite{gill1995applications} which bounds from below the MSE of any estimator of $\theta$ by the inverse of the expected value of the aforementioned Fisher information plus $I_0$. The details are given in the Appendix.\\

Next, we present an adaptive estimation scheme which attains ARE of $\pi/2$. 
\subsection{Asymptotically optimal estimator}
Consider the following estimator $\widehat{\theta}_n$ for $\theta$:  set 
\begin{equation}
\label{eq:sgd_alg}
\theta_n = \theta_{n-1} +  \gamma_n \sgn (X_n - \theta_n), \quad n = 1,2,\ldots,
\end{equation}
where $\left\{\gamma_n \right\}_{n=1}^\infty$ is any strictly positive sequence satisfying 
\begin{enumerate}
\item[(i)] $\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n)$ \\
\item[(ii)] $\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty$ 
for some $0< \lambda \leq 1$.
\end{enumerate}
(e.g. $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). The $n$th step estimation is defined by 
\begin{equation} \label{eq:sgd_est}
\widehat{\theta}_n =  \frac{1}{n} \sum_{i=1}^n  \theta_i. 
\end{equation}

For the estimator defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg} 
we have the following results:
\begin{thm} \label{thm:sgd}
The sequence $\widehat{\theta}_n$ of \eqref{eq:sgd_est} satisfies
\[
\sqrt{n} \left( \widehat{\theta}_n - \theta \right) \overset{d}{\rightarrow} \mathcal N \left(0,  \pi \sigma^2 /2 \right).
\]
\end{thm}

@@

In other words, Thm.~\ref{thm:sgd} implies that the estimator $\widehat{\theta}_n$, defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg}, attains the minimal asymptotic efficiency as established by Thm.~\ref{thm:adpative_lower_bound}.\par
Note that $\theta_0$ is not explicitly defined 
in equation \eqref{eq:sgd_est}. While a reasonable initialization is $\theta_0 = \mathbb E [\theta]$,  Thm.~\ref{thm:sgd} implies that the asymptotic behavior of the estimator is indifferent to this initialization. Thus, the optimal efficiency is attained regardless of the prior distribution on $\theta$ or the size of the parameter space $\Theta$. Nevertheless, the bound in Thm.~\ref{thm:adpative_lower_bound} suggests that the non-asymptotic estimation error can be greatly reduced whenever the location information $I_0$ is large. In contrast, the one-step optimal scheme presented in the following subsection exploits the prior information on $\theta$ provided by $\pi(\theta)$.

\subsection{One-step optimal estimation}
We now consider an estimation scheme that posses the property of \emph{one-step optimality}: at each step $i$, the $i$th encoder designs the detection region $M_i^{-1}(1)$ such that the MSE given $M^i$ is minimal. In other word, this scheme designs the messages in a greedy manner, such that the MSE at step $i$ is minimal given the current state of the estimation described by $M^{i-1}$. \\

The following theorem determine the structure of the message that minimizes the next step MSE:
\begin{thm}[optimal one-step estimation] \label{thm:opt_one_step}
Let $\pi(\theta)$ be an absolutely continuous log-concave probability distribution. Given a sample $X$ from the distribution $\mathcal N(\theta, \sigma^2)$, define 
\begin{equation}
\label{eq:adaptive_main_message}
M = \sgn(X - \tau),
\end{equation}
where $\tau$ satisfies the equation
\begin{equation}
 \label{eq:fixed_point}
 \tau = \frac{m^-(\tau) + m^+(\tau)}{2},
\end{equation}
with
\begin{align*}
m^-(\tau)  & = \frac{\int_{-\infty}^{\tau} \theta \pi(d\theta) }{\int_{-\infty}^{\tau} \pi(d\theta)} ,\\
m^+(\tau) & = \frac{\int_{\tau}^\infty \theta \pi(d\theta) }{\int_{\tau}^\infty \pi(d\theta)} .
\end{align*}
Then for any estimator $\widehat{\theta}$ which is a function of $M'(X) \in \{-1,1\}$, we have
\begin{equation}
\label{eq:opt_cond}
\mathbb E \left(\theta-\widehat{\theta}(M')\right)^2 \geq  \mathbb E \left(\theta- \mathbb E[\theta|M]\right)^2,
\end{equation}
\end{thm}

\begin{proof}
The proof is completed by the following two lemmas, proofs of which can be found in the Appendix:
\begin{lem} \label{lem:unique}
Let $f(x)$ be a log-concave probability density function. Then the equation 
\begin{equation}
\label{eq:lem_fixed_point}
2x = \frac{\int_x^\infty uf(u)du}{\int_x^\infty f(u)du} + \frac{\int_{-\infty}^x uf(u)du}{\int_{-\infty}^x f(u)du} 
\end{equation}
has a unique solution.
\end{lem}
\begin{lem} \label{lem:adaptive}
Let $U$ be an absolutely continuous random variable with
pdf $P(du)$. Then the one-bit message $M^\star\in \{-1,1\}$ that minimizes
\[
\int \left( u - \mathbb E[U|M(u)]  \right)^2 P(du)
\]
is given by
\[
M^\star  =  \sgn(U - \tau),
\]
where $\tau$ is the unique solution to
 \[
2 \tau = \frac{\int_{\tau}^\infty u P(du)} {\int_{\tau}^\infty P(du)} + \frac{\int_{-\infty}^{\tau} u P(du)}{\int_{-\infty}^{\tau} P(du)}.
\]
\end{lem}
\end{proof}

\begin{rem}
The value of the optimal threshold as given in Theorem~\ref{thm:opt_one_step} is different than the one given in \cite[Eq. 5]{1619423} for apparently the same problem. Indeed, it seems like Equation 4 there is erroneous. 
\end{rem}

Thm.~\ref{thm:opt_one_step} suggests the following adaptive encoding and estimation scheme: 
\begin{itemize}
\item Initialization: set $P_0(t) = \pi(\theta)$.
\item For $n\geq 1$:
\begin{enumerate}
\item Update the prior as
\begin{align}
P_n(t) = & P(\theta=t |M^n) \\
& = \frac{ P\left( \theta=t | M^{n-1} \right) P(M_n | \theta = t , M^{n-1})  } { P(M_n | M^{n-1} )} \nonumber \\ 
& = \alpha_n  P_{n-1}(t) \Phi\left(M_n \frac{ t - \tau_{n-1} }{\sigma} \right), \label{eq:density_update}
\end{align}
where $\alpha_n$ is a normalization coefficient that equals to
\[
\alpha_n = \left(\int_{\mathbb R} P_{n-1}(t) \Phi\left(M_n \frac{t- \tau_{n-1} }{\sigma} \right)  dt \right)^{-1}. 
\]
\item The $n$th estimate for $\theta$ is the conditional expectation of $\theta$ given $M^n$, namely
\begin{equation}
\theta_n = \mathbb E \left[ \theta| M^n\right] = \int_{-\infty}^\infty t P_n(t) dt. \label{eq:estimator_update}
\end{equation}
\item Solve equation \eqref{eq:fixed_point} with the updated prior $P_n(t)$ instead of $P(d\theta)$. Note that since the standard normal cdf $\Phi(x)$ is log-concave, the updated prior $P_n(t)$ remains log-concave and thus a unique solution to \eqref{eq:fixed_point} is guaranteed by Lem.~\ref{lem:unique}. 
\item Update the $(n+1)$th message as
\begin{equation}\label{eq:message_update}
M_{n+1} = \sgn(X_{n+1}-\tau_n)
\end{equation}
\end{enumerate}
\end{itemize}
Since equation \eqref{eq:fixed_point} has no analytic solution in general, it is hard to derive the asymptotic behavior of the estimator defined by \eqref{eq:estimator_update} and \eqref{eq:message_update}. We conjecture, however, that it attains the asymptotic relative efficiency of $\sigma^2\pi/2$, as can be observed from the numerical simulation illustrated in Fig.~\ref{fig:adaptive_error}. Also shown in Fig.~\ref{fig:adaptive_error} are the normalized MSE of the asymptotically optimal estimator defined by \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, as well as the MSE achieved by the sample mean for the same sample realization.

\begin{figure}
\begin{center}
\begin{tikzpicture}
\node at (0,0) {\includegraphics[scale=0.4]{one_bit_adpative}};
%\node[color = blue,  minimum width=1cm, align = left] at (1.5,1.5) {assymptotically optimal \eqref{eq:sgd_est}};
%\node[color = red, minimum width=1cm, align = left] at (1.5,0.55) {one-step optimal \eqref{eq:message_update}} ;
\node[rotate = 90, scale = 0.7] at (-3.8,0) {$n \mathbb E \left(\widehat{\theta}_n - \theta \right)^2$};
\node[scale = 0.7] at (0,-2.4) {$n$};
%\draw[dashed] (-3.5,1.18) node[left, scale = 0.7] {$\frac{\pi}{2}$}-- (-3,1.18);
\end{tikzpicture}
\caption{Normalized empirical risk $n\left(\widehat{\theta}_n-\theta\right)^2$ versus number of samples $n$ for $500$ Monte Carlo trials. In each trial, $\theta$ is chosen uniformly in the interval $(-3,3)$. \label{fig:adaptive_error}  }
\end{center}
\end{figure}

@@

\subsection{Threshold Detection}
We assume that each message is of the form
\[
M_i = \sgn(t_i - X_i) = \begin{cases} 1 & X_i< t_i, \\
-1 & X_i > t_i,
\end{cases}  
\]
where $t_i\in\mathbb R$ is the \emph{threshold} of the $i$th detector applied by the encoder on it sample $X_i$. In order to characterize the asymptotic behavior of estimation from these sequence of messages, we consider the \emph{density} of $n$ threshold values defined as:
\[
\lambda_n([a,b]) = \frac{1}{n} \left| \mathcal T \cap [a,b] \right|,
\]
where $\mathcal T  = \left\{t_n, n \in \mathbb N \right\}$ is the collection of all threshold values. We further assume that $\lambda_n$ converges (weakly) to a probability measure $\lambda(t)$ on $\mathbb R$.  \par
In order to estimate $\theta$ from the messages $M^n$, we consider the maximum likelihood (ML) estimator. The log-likelihood function of $\theta$ is given by
\[
l(M^n  |\theta ) = \sigma  \sum_{i=1}^n \log \left( \Phi \left( M_i \frac{t_i - \theta}{\sigma} \right) \right). 
\]
Since $\Phi(x)$ is a log concave function, the log-likelihood function has a unique maximizer $\widehat{\theta}_n$ which is the ML estimator. Specifically, the ML estimator $\widehat{\theta}_{ML}$ is the unique root of 
\[
\sum_{i=1}^n M_i \frac{\phi \left( \frac{t_i-\theta}{\sigma}\right) }{\Phi \left(M_ii \frac{t_i-\theta}{\sigma}\right) }
\]

Next, we show that under the assumptions above the sequence of messages $M^n$ defines a local asymptotic normal (LAN) family of probability distributions. Therefore, the MSE in estimating $\theta$ from the messages $M^n$ satisfies a local asymptotic minimax property with respect to the \emph{precision} parameter of this LAN family \cite{van2000asymptotic}. Moreover, we conclude the ML estimator is locally asymptotic minimax.

@@

\begin{rem}
The condition that $\lambda_n$ converges weakly to $\lambda$ can be relaxed to the condition
\[
\frac{1}{n} \sum_{i=1}^n  \eta \left( \frac{t_i-\theta}{\sigma} \right)  \overset{n \rightarrow \infty}\longrightarrow \int \eta \left( \frac{t-\theta}{\sigma} \right) \lambda(dt)
\]
(pointwise for any $\theta \in \Theta$). 
\end{rem}
